### 2024-10-03 MetadataS1/MetadataS1.tex
Revision of the entire document and minor changes

\documentclass[a4paper,twoside,12pt]{article}
%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}
\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{parskip}
\usepackage{enumitem}
%% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx} %%For loading graphic files
\usepackage{subcaption}
\usepackage{float}
\usepackage{booktabs} %% for top/midrule
\usepackage{textcomp}
\usepackage[table,xcdraw, dvipsnames]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{makecell}
\usepackage{pdflscape}
%bibliography
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0} % Removes the border
}
\usepackage{geometry}
%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\onehalfspacing       %% 1,5-spacing
\usepackage{placeins}
\usepackage{lineno}
\linenumbers
\usepackage{titlesec}
\titlespacing{\section}{0pt}{*4}{*1.5} 
\titlespacing{\subsection}{20pt}{*4}{*1.5}
\titlespacing{\subsubsection}{40pt}{*4}{*1.5} 
\titlespacing{\paragraph}{60pt}{*4}{*1.5}
\titlespacing{\subparagraph}{80pt}{*4}{*1.5}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

@@

\subsection*{Introduction}
 The composition of ecological communities, defined as the abundance of each species within a given community, is fundamental for understanding patterns and processes in community ecology. Variations in community composition can help to detect spatial patterns linked to environmental variations \citep{kemp1990}, assess temporal trends of different groups of species after disturbances \citep{philippi1998, magurran2007}, and understand food web structures \citep{cohen2003}. Additionally, community composition is essential for modeling the dynamics of ecological communities. Dynamic community modelling allows addressing important issues and questions in ecology, such as: determining the relative strength of top-down versus bottom-up forces in communities \citep{krebs2003,legagneux2014}, assessing the ecological resilience of communities under climate change \citep{griffith2019} and evaluating the cascading effects of invasive species in food web \citep{david2017, goto2020}. Dynamic community modelling can also be applied to address practical challenges, including fishery management \citep{plaganyi2007} and the planning of protected areas \citep{okey2004, dahood2020}. 
Modeling food webs requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community. However, accurately determining the abundance of all species is rarely achievable. Consequently, empirical community models often reduce taxonomic resolution by grouping species into large functional or taxonomic categories. Additionally, food webs consist of species with varying body sizes depending on their trophic level, with top-level species often being highly mobile and having large home ranges \citep{mccann2005}. Therefore, community models must use landscape-wide estimates of species abundance to accurately represent trophic fluxes. Due to these constraints, empirical datasets with high taxonomic resolution that cover entire communities at broad spatial and temporal scales are rare and often include incomplete or rough estimates.

@@

Arctic environments are highly valuable systems for studying community patterns and processes due to their relatively low species richness \citep{payer2013, legagneux2014}. However, logistical challenges in the Arctic limit the number of long-term biodiversity monitoring programs. Hence, the small number of Arctic communities with long-term monitoring serve as highly valuable sites for holistic and empirical community studies. Datasets on terrestrial communities are notably scarce, and this scarcity extends to Arctic communities as well \citep{ims2013}. 

@@

\begin{figure}[H]
\includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
\caption{Synthetic vertebrate food web of the south plain of Bylot Island.}
\label{figure:food_web}
\subsection*{Objective}
Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments. Additionally, we aim to provide the average body mass for each species in the community, enabling the conversion of abundances into biomasses.
\section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
    \section*{A. Data set identity} Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut
    \section*{B. Data set identification codes}
     \texttt{BYLOT-species\_taxonomy.csv}\\
     \texttt{BYLOT-species\_abundance.csv}\\
     \texttt{BYLOT-mean\_species\_abundance.csv}\\
     \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
     \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

@@

\section*{\textcolor{Blue}{Class II. Research origin descriptors}}
    \section*{A. Overall project description}
    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial vertebrate communities
        \subsection*{2. Originators:}   
        \textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
                \textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada

@@

        \subsection*{4. Objectives}
       i) Understand the factors that shape the structure and drive the dynamics of Arctic terrestrial vertebrate communities.\\
       ii) Predict the effects of current global environmental changes on the structure and dynamics of Arctic terrestrial vertebrate communities.

     \subsection*{5. Abstract}
        Arctic terrestrial vertebrate communities present low species richness, making those relatively simple communities ideal for studying ecological patterns and dynamics in terrestrial environments. Although, they present complex networks of interacting species, extreme seasonal changes of environmental conditions and a large portion of migratory species, which make challenging the identification of the key factors that shape their structure and dynamics. In the face of rapid global environmental changes, it is crucial to have a comprehensive understanding of the key processes shaping Arctic terrestrial communities structure and dynamics in order to predict how global changes will impact them in the future. Our research emphasizes long-term biodiversity monitoring, a community-wide perspective and food web modeling to achieve this understanding.

@@

    \section*{B. Specific subproject description}
        \subsection*{1. Site description}
                        \subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic. An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:table_species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:table_species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.
                        \input{tables/table_species_name_strategy.tex}
%label-> table:table_species_name_strategy
\newpage
                \subsubsection*{b. Geography} Our 463 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
                \subsubsection*{c. Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
                \subsubsection*{d. Geology} See \citet{klassen1993} for a complete and detailed description of the geology of the study area.
                \subsubsection*{e. Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
                \subsubsection*{f. Site history} See \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
                \subsubsection*{g. Climate} The mean temperature in July is 6°C, and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}. The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds. \citep{gauthier2024a}.
 \newpage

@@

   \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada. The outline of the snow goose colony area is presented with white dash; we used the outline of the colony in 2017 since it represents an average colony area (74 km\textsuperscript{2}).}

@@

  \caption{Intensive nests searching plots (8 km\textsuperscript{2} and 2 km\textsuperscript{2}) located in the Qarlikturvik valley.}

@@

        Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area. This sampling aimed to find all nests of the cackling goose and the glaucous gull. Nest locations of other large wetland-nesting species, including the tundra swan, the red-throated loon and the Pacific loon, were also noted, as these species nest in the same habitat \citep{duchesne2021,gauthier2024a}. Each year, all known or potential nesting sites were revisited. Observers detected nests by walking and scanning around ponds and lakeshores to identify any active nesting sites. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. Most of them (geese, swans and gulls) can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species given the open landscape.\\

        \item[] \textit{\textbf{Snow goose}}\\
        Snow geese nest in a large colony in the study area (\textbf{Figure \ref{figure:zones}}), but also in small aggregations distributed on the island, especially in years when snowy owls are nesting \citep{lepage1996,reed2002}. Since 1994, goose nests were systematically monitored on a 0.24 km\textsuperscript{2} wetland at the center of the colony. Since 1999, nests were also systematically monitored on a variable number of plots, measuring 0.01 km\textsuperscript{2} in wetland habitat and 0.04 km\textsuperscript{2} in mesic habitat, randomly distributed throughout the goose colony \citep{gauthier2020goose}. The total area covered by the randomly distributed plots averaged 0.79 \textpm{0.37} km\textsuperscript{2} per year. From 2010 onwards, except in 2020 and 2021, we traced the approximate boundary of the goose colony using a GPS receiver aboard a helicopter flying along the colony border \citep{duchesne2021}.

@@

        \item[] \textit{\textbf{Snowy owl}} \\
        Snowy owls predominantly nest in habitats similar to other raptors, favoring ridges in mountainous or hilly regions, although they can occasionally be found nesting on mounds in lowland areas \citep{seyer2020}. Since 1996, searches for snowy owl nests have been conducted concurrently with searches for other raptor nests in the Black and Southern plateaus, as well as during searches for jaeger nests on the southern side of the glacial river in the Qarlikturvik Valley. Additionally, since 2012, nests have been recorded across the entire study area by scanning the landscape from hills and ridges during the nesting period \citep{duchesne2021}. Given that snowy owls nest on elevated mounds, exhibit contrasting colors with the landscape, emit alarm calls, and display defensive behaviors, active nesting sites have a high probability of detection.\\

        \item[] \textit{\textbf{Long-tailed jaeger and parasitic jaeger}} \\
        Since 2004, observers have walked parallel transects spaced 400 meters apart, covering the entire southern side of the glacial river in the Qarlikturvik Valley (33 km\textsuperscript{2}; \textbf{Figure \ref{figure:qarlikturvik_valley}}), during the nesting period. The aim of those transects was to record nests of long-tailed jaegers, parasitic jaegers, and sandhill cranes. Observers listened for alarm calls to detect territorial birds, and then located nests by observing the birds returning to their nests from elevated vantage points. We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability. We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).\\

        \item[] \textit{\textbf{Common-ringed plover}} \\
        Between 2015 and 2019, observers conducted surveys of the primary nesting areas of the common-ringed plover. The survey involved walking in stony and sandy shores and gravel bars with scarce vegetation along rivers. Nests were found by detecting individuals exhibiting reproductive behaviors, such as incubation, alarm calls, or distraction displays. The sampling effort was particularly intensive between 2015 and 2017. Small areas along the coast or on the banks of smaller rivers that could potentially serve as nesting sites may have been overlooked.\\

        \item[] \textit{\textbf{Lapland longspur and Baird's sandpiper}}\\
        Since 2005, nests of passerines and sandpipers have been extensively monitored across an 8 km\textsuperscript{2} (4x2 km) area in the Qarlikturvik valley. We considered the sampling to be most systematic within a core 2 km\textsuperscript{2} (2x1 km) plot in this area (\textbf{Figure \ref{figure:qarlikturvik_valley}}). We excluded relatively large water bodies (0.26 km\textsuperscript{2}) to calculate nest density in the plot due to the presence of a large lake, which leaves an area of 1.74 km\textsuperscript{2} available for nesting. An observer conducted systematic searches of this plot during the entire breeding season to locate and monitor as many passerine and shorebird nests as possible. Assuming the observer can detect all nests within a 5 or 10 meter radius, analysis of daily GPS tracks shows that the observer covered a minimum area of 0.72 \textpm{0.12} (5 m) or 1.09 \textpm{0.17} km\textsuperscript{2} (10 m) of the core area annually (n= 3 years). Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically.\\

        \item[] \textit{\textbf{Opportunistic nest monitoring}}\\
        Since 2005, we also noted the nest location of any other bird species encountered opportunistically during travel or while carrying out the protocols for the previously described species. The sampling was particularly intensive in the defined 8 km\textsuperscript{2} (2x4 km plot) area in the Qarlikturvik valley. The accuracy of nest monitoring in this plot thus depends on the species detection probability. We are confident to obtain a realistic order of magnitude for the number of nests present for relatively large bodied species in this area (i.e., sandhill crane, rock ptarmigan, long-tailed duck and king eider). Additionally, starting in 2009, a significant effort has been made each year, though not systematically, to visit known nesting territories of parasitic jaegers throughout the study area.
\input{tables/table_interannual_nest_density_variation.tex}
%label-> table:table_interannual_nest_density_variation

@@

Since 2007, observers have recorded all vertebrate species observed opportunistically during field work and tallied the total number of individuals at the end of each day \citep{gauthier2020daily, gauthier2024a}. The number of hours spent in the field served as a proxy for the sampling effort. We used the number of individuals observed per hour spent in the field calculated by \citet{gauthier2024a} as an index of relative abundance for each species. However, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands. \\

@@

                        \subsection*{2. Size} 
                        a. 35 records, not including header row (4.3 kB)\\
                        b. 271 records, not including header row  (34.5 kB)\\
                        c. 53 records, not including header row  (3.5 kB)\\
                        d. 18 records, not including header row  (961 B)

>>>>

\documentclass[a4paper,twoside,12pt]{article}
%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}
\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{parskip}
\setlength{\parskip}{12pt}

%% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx} %%For loading graphic files
\usepackage{subcaption}
\usepackage{float}
\usepackage{booktabs} %% for top/midrule
\usepackage{textcomp}
\usepackage[table,xcdraw, dvipsnames]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{makecell}
\usepackage{pdflscape}
%bibliography
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0} % Removes the border
}
\usepackage{geometry}
%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\onehalfspacing       %% 1,5-spacing
\usepackage{placeins}
\usepackage{lineno}
\linenumbers
\usepackage{titlesec}
\titlespacing{\section}{0pt}{*4}{*1.5} 
\titlespacing{\subsection}{20pt}{*4}{*1.5}
\titlespacing{\subsubsection}{40pt}{*4}{*1.5} 
\titlespacing{\paragraph}{60pt}{*4}{*1.5}
\titlespacing{\subparagraph}{80pt}{*4}{*1.5}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

@@

\subsection*{Introduction}
 The composition of ecological communities, defined as the abundance of each species within a given community, is fundamental for understanding patterns and processes in community ecology. Variations in community composition can help to detect spatial patterns linked to environmental variations \citep{kemp1990}, assess temporal trends of different groups of species after disturbances \citep{philippi1998, magurran2007}, and understand food web structures \citep{cohen2003}. Additionally, community composition is essential for modeling the dynamics of ecological communities. Dynamic community modelling allows addressing important issues and questions in ecology, such as: determining the relative strength of top-down versus bottom-up forces in communities \citep{krebs2003,legagneux2014}, assessing the ecological resilience of communities under climate change \citep{griffith2019} and evaluating the cascading effects of invasive species in food web \citep{david2017, goto2020}. Dynamic community modelling can also be applied to address practical challenges, including fishery management \citep{plaganyi2007} and the planning of protected areas \citep{okey2004, dahood2020}.

Modeling food webs requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community. However, determining the abundance of all species within a community is rarely achievable. Consequently, empirical community models often reduce taxonomic resolution by grouping species into large functional or taxonomic categories. Additionally, food webs consist of species with varying body sizes depending on their trophic level, with top-level species often being highly mobile and having large home ranges \citep{mccann2005}. Therefore, community models must use landscape-wide estimates of species abundance to accurately represent trophic fluxes. Due to these constraints, empirical datasets with high taxonomic resolution that cover entire communities at broad spatial and temporal scales are rare and often include incomplete or rough estimates.

@@

Arctic environments are highly valuable systems for studying community structure and dynamics due to their relatively low species richness \citep{payer2013, legagneux2014}. However, logistical challenges in the Arctic limit the number of long-term biodiversity monitoring programs. Hence, the small number of Arctic communities with long-term monitoring serve as highly valuable sites for holistic and empirical community studies. Datasets on terrestrial communities are notably scarce, and this scarcity extends to Arctic communities as well \citep{ims2013}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
\caption{Synthetic vertebrate food web of the south plain of Bylot Island.}
\label{figure:food_web}
\end{figure}

\subsection*{Objective}
Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. We focus on adults, except for lemmings for which we have not distinguished between juveniles and adults. Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments. Additionally, we aim to provide the average body mass for each species in the community, enabling the conversion of abundances into biomasses.
\newpage
\section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
    \section*{A. Data set identity} Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut
    \section*{B. Data set identification codes}
     \texttt{BYLOT-species\_taxonomy.csv}\\
     \texttt{BYLOT-species\_abundance.csv}\\
     \texttt{BYLOT-species\_body\_mass.csv}\\
     \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

@@

\section*{\textcolor{Blue}{Class II. Research origin descriptors}}
    \section*{A. Overall project description}
    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial communities
        \subsection*{2. Originators}   
        \textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
                \textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada

@@

        \subsection*{4. Objectives}
       i) Understand the factors that shape the structure and drive the dynamics of Arctic terrestrial communities.\\
       ii) Predict the effects of current global environmental changes on the structure and dynamics of Arctic terrestrial communities.

     \subsection*{5. Abstract}
      Arctic terrestrial communities, characterized by relatively low species richness, offer unique opportunities for studying ecological patterns and dynamics in simplified systems. Despite their relative simplicity, these ecosystems feature complex species interactions, extreme seasonal environmental changes, and a significant proportion of migratory species, making it difficult to identify the key factors shaping their structure and dynamics. As global environmental changes accelerate, it is essential to understand the processes driving these communities to eventually predict future impacts on Arctic ecosystems. Our research combines long-term biodiversity monitoring, a community-wide approach, and food web modeling to address these challenges.

@@

    \section*{B. Specific subproject description}
        \subsection*{1. Site description}
                        \subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic. An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:table_species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:table_species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.
                        \input{tables/table_species_name_strategy.tex}
%label-> table:table_species_name_strategy
\newpage
                \subsubsection*{b. Geography} Our 389 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
                \subsubsection*{c. Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
                \subsubsection*{d. Geology} See \citet{klassen1993} for a detailed description of the geology of the study area.
                \subsubsection*{e. Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Louis-Pierre Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
                \subsubsection*{f. Site history} See \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
                \subsubsection*{g. Climate} The mean annual air temperature since 1995 is -14.4°C, with mean seasonal temperature of 4.7°C in summer (June to August), -11°C in fall (September to November), -32.4°C in winter (December to February) and -19.4°C in spring (March to May; \citet{bylotclimate}). The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds \citep{gauthier2024a}. In summer, the study area received on average 77.5 mm of precipitation \citep{bylotclimate}. Additionnally, 102 days are frost-free annually on average  \citep{bylotclimate} and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}.
\newpage

@@

   \caption{Map of the different zones (colored polygons) of the 389 km\textsuperscript{2} study area located on the south plain of Bylot Island, Nunavut Canada. The perimeter of the snow goose colony is delineated by white dashes; we highlighted the perimeter in 2017 since it represents the average colony area (74 km\textsuperscript{2}).}

@@

  \caption{Intensive nests searching plots within the Qarlikturvik valley.}

@@

        Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik valley, and since 2017, in other zones of the study area. This sampling aimed to find all nests of the cackling goose and the glaucous gull. Nest locations of other large wetland-nesting species, including the tundra swan, the red-throated loon and the Pacific loon, were also noted, as these species nest in the same habitat \citep{duchesne2021,gauthier2024a}. Each year, all known or potential nesting sites were revisited. Observers detected nests by walking and scanning around ponds and lakeshores to identify any active nesting sites. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. Most of them (geese, swans and gulls) can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species given the open landscape.

        \item[] \textit{\textbf{Snow goose}}\\
        Snow geese nest in a large colony in the study area (\textbf{Figure \ref{figure:zones}}), but also in small aggregations distributed on the island, especially in years when snowy owls are nesting \citep{lepage1996,reed2002}. Since 1994, goose nests were systematically monitored on a 0.24 km\textsuperscript{2} wetland at the center of the colony. Since 1999, nests were also systematically monitored on a variable number of plots, measuring 0.01 km\textsuperscript{2} in wetland habitat and 0.04 km\textsuperscript{2} in mesic habitat, randomly distributed throughout the goose colony \citep{gauthier2020goose}. The total area covered by the randomly distributed plots averaged 0.79 \textpm{0.37} km\textsuperscript{2} per year. From 2010 onwards, except in 2020 and 2021, we opportunistically traced sections of the approximate boundary of the goose colony using a GPS receiver aboard a helicopter, taking advantage of regular flights across the study area whenever the flight path passed over the colony border \citep{duchesne2021}.

@@

        \item[] \textit{\textbf{Snowy owl}} \\
        Snowy owls predominantly nest in habitats similar to other raptors, favoring ridges in mountainous or hilly regions, although they can occasionally be found nesting on mounds in lowland areas \citep{seyer2020}. Since 1996, searches for snowy owl nests have been conducted concurrently with searches for other raptor nests in the Black and Southern plateaus, as well as during searches for jaeger nests on the southern side of the glacial river in the Qarlikturvik Valley. Additionally, since 2012, nests have been recorded across the entire study area by scanning the landscape from hills and ridges during the nesting period \citep{duchesne2021}. Given that snowy owls nest on elevated mounds, exhibit contrasting colors with the landscape, emit alarm calls, and display defensive behaviors, active nesting sites have a high probability of detection.

        \item[] \textit{\textbf{Long-tailed jaeger and parasitic jaeger}} \\
        Since 2004, observers have walked parallel transects spaced 400 meters apart, covering the entire southern side of the glacial river in the Qarlikturvik Valley (33 km\textsuperscript{2}; \textbf{Figure \ref{figure:qarlikturvik_valley}}), during the nesting period. The aim of those transects was to record nests of long-tailed jaegers, parasitic jaegers, and sandhill cranes. Observers listened for alarm calls to detect territorial birds, and then located nests by observing the birds returning to their nests from elevated vantage points. We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability. We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).

        \item[] \textit{\textbf{Common-ringed plover}} \\
        Between 2015 and 2019, observers conducted surveys of the primary nesting areas of the common-ringed plover. The survey involved walking in stony and sandy shores and gravel bars with scarce vegetation along rivers. Nests were found by detecting individuals exhibiting reproductive behaviors, such as incubation, alarm calls, or distraction displays. The sampling effort was particularly intensive between 2015 and 2017. Small areas along the coast or on the banks of smaller rivers that could potentially serve as nesting sites may have been overlooked.

        \item[] \textit{\textbf{Lapland longspur and Baird's sandpiper}}\\
        Since 2005, nests of passerines and sandpipers have been extensively monitored across an 8 km\textsuperscript{2} (4x2 km) area in the Qarlikturvik valley. We considered the sampling to be most systematic within a core 2 km\textsuperscript{2} (2x1 km) plot in this area (\textbf{Figure \ref{figure:qarlikturvik_valley}}). We excluded relatively large water bodies (0.26 km\textsuperscript{2}) to calculate nest density in the plot due to the presence of a large lake, which leaves an area of 1.74 km\textsuperscript{2} available for nesting. An observer conducted systematic searches of this plot during the entire breeding season to locate and monitor as many passerine and shorebird nests as possible. Assuming the observer can detect all nests within a 5 or 10 meter radius, analysis of daily GPS tracks shows that the observer covered a minimum area of 0.72 \textpm{0.12} (5 m) or 1.09 \textpm{0.17} km\textsuperscript{2} (10 m) of the core area annually (n= 3 years). Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically.

        \item[] \textit{\textbf{Opportunistic nest monitoring}}\\
        Since 2005, we also noted the nest location of any other bird species encountered opportunistically during travel or while carrying out the protocols for the previously described species. The sampling was particularly intensive in the defined 8 km\textsuperscript{2} area in the Qarlikturvik valley. The accuracy of nest monitoring in this plot thus depends on the species detection probability. We are confident to obtain a realistic order of magnitude for the number of nests present for relatively large bodied species in this area (i.e., sandhill crane, rock ptarmigan, long-tailed duck and king eider). Additionally, starting in 2009, a significant effort has been made each year, though not systematically, to visit known nesting territories of parasitic jaegers throughout the study area.
        \end{enumerate}

@@

Since 2007, observers have recorded all vertebrate species observed opportunistically during field work and tallied the total number of individuals at the end of each day \citep{gauthier2020daily, gauthier2024a}. The number of hours spent in the field served as a proxy for the sampling effort. We used the number of individuals observed per hour spent in the field calculated by \citet{gauthier2024a} as an index of relative abundance for each species. Moreover, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands. \\

@@

                        \subsection*{2. Size} 
                        a. 35 records, not including header row (4.3 kB)\\
                        b. 271 records, not including header row  (33.0 kB)\\
                        c. 53 records, not including header row  (3.7 kB)\\
                        d. 18 records, not including header row  (961 B)
